Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-10-02 tikz_images/proof.tex
Added cryptography proof summary tikz diagram - changed text a tiny bit to reference it nicely Moved tikz image files to their own folder - nicer to keep the imported pdfs and the itkz picture (which aren't imported) seperate

% autosam.tex
% Annotated sample file for the preparation of LaTeX files
% for the final versions of papers submitted to or accepted for 
% publication in AUTOMATICA.

% See also the Information for Authors.

% Make sure that the zip file that you send contains all the 
% files, including the files for the figures and the bib file.

% Output produced with the elsart style file does not imitate the
% AUTOMATICA style. The style file is generic for all Elsevier
% journals and the output is laid out for easy copy editing. The
% final document is produced from the source file in the
% AUTOMATICA style at Elsevier.

% You may use the style file autart.cls to obtain a two-column 
% document (see below) that more or less imitates the printed 
% Automatica style. This may helpful to improve the formatting 
% of the equations, tables and figures, and also serves to check 
% whether the paper satisfies the length requirements.

% Please note: Authors must not create their own macros.

% For further information regarding the preparation of LaTeX files 
% for Elsevier, please refer to the "Full Instructions to Authors" 
% from Elsevier's anonymous ftp server on ftp.elsevier.nl in the
% directory pub/styles, or from the internet (CTAN sites) on
% ftp.shsu.edu, ftp.dante.de and ftp.tex.ac.uk in the directory
% tex-archive/macros/latex/contrib/supported/elsevier.

% The use of LaTeX2e is preferred.
%\documentclass{elsart}

% Enable this line and disable the 
% preceding line to obtain a two-column 
% document whose style resembles the
% printed Automatica style.
\documentclass[twocolumn]{autart}

% Include this line if your 
% document contains figures,
\usepackage{graphicx}
% or this line, depending on which
% you prefer.
%\usepackage[dvips]{epsfig}

% User imported packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}

% Local packages
\usepackage{ISASmacros/isasmathmacros}

% Algorithm settings
\algnewcommand{\LineComment}[1]{\State \(\triangleright\) #1}

% Tikz settings
\usetikzlibrary{math}

% Tikz reused pics
\tikzset{
   plane/.pic = {
    \draw[fill]  plot[smooth, tension=0.6] coordinates {
        (-0.65,-0.9) 
        (-0.6,-0.85) 
        (-0.4,-0.75) 
        (-0.25,-0.65) 
        (-0.15,-0.5) 
        (-0.12,-0.3) 
        (-0.1,-0.1) 
        (0,0) 
        (0.1,-0.1) 
        (0.12,-0.3) 
        (0.15,-0.5) 
        (0.25,-0.65) 
        (0.4,-0.75) 
        (0.6,-0.85) 
        (0.65,-0.9)
        } -- plot[smooth, tension=0.6] coordinates {
        (0.65,-0.9) 
        (0.15,-0.91)
        (0.35,-1.1) 
        (0.37,-1.15)
        } -- plot[smooth, tension=0.6] coordinates {
        (0.37,-1.15)
        (0,-1.12) 
        (-0.37,-1.15) 
        } -- plot[smooth, tension=0.6] coordinates {
        (-0.37,-1.15)
        (-0.35,-1.1) 
        (-0.15,-0.91) 
        (-0.65,-0.9) 
        } -- cycle;
   }
}

\begin{document}

\begin{frontmatter}
% Running title for regular 
% papers but only if the title  
% is over 5 words. Running title 
% is not shown in output.
%\runtitle{Insert a suggested running title}

% Title, preferably not more than 10 words.
\title{Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}}

\thanks[footnoteinfo]{This paper was not presented at any IFAC 
meeting. Corresponding author Uwe~D.~Hanebeck. Tel. +49--721--608--43909.}

% Add the 
% e-mail address 
% (ead) as shown
\author[ISAS]{Marko Ristic}\ead{marko.ristic@kit.edu},
\author[ISAS]{Benjamin Noack}\ead{noack@kit.edu},
\author[ISAS]{Uwe D. Hanebeck}\ead{uwe.hanebeck@kit.edu}

% Please supply
\address[ISAS]{Intelligent Sensor-Actuator-Systems Laboratory, Institute for Anthropomatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}

% Five to ten keywords,  
% chosen from the IFAC 
% keyword list or with the 
% help of the Automatica 
% keyword wizard
\begin{keyword}
System state estimation; Data privacy; Sensor fusion; Kalman filters.
\end{keyword}

% Abstract of not more than 200 words.
\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor information or navigator location during localisation protocols and fail to preserve participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common, model-based, non-linear measurement, localisation methods and forfeit broad applicability. We define a cryptographically secure linear-combination aggregation scheme which we apply to the Extended Kalman Filter with range-sensor measurements, and show that navigator location, sensor locations and sensor measurements can remain private during navigation. The security requirements, leakage, and cryptographic proof are given for the private filter and aggregation scheme, and simulations of the filter are used to evaluate the accuracy and performance of the method. Our approach defines a novel, computationally plausible and cryptographically private, model-based localisation filter with direct application to environments where nodes may not be fully trusted and data is considered sensitive.
\end{abstract}

\end{frontmatter}

% 
% 8888888          888                    
%   888            888                    
%   888            888                    
%   888   88888b.  888888 888d888 .d88b.  
%   888   888 "88b 888    888P"  d88""88b 
%   888   888  888 888    888    888  888 
%   888   888  888 Y88b.  888    Y88..88P 
% 8888888 888  888  "Y888 888     "Y88P"  
%                                         
%                                         
%                                         
%

\section{Introduction}

--

Introduce localisation, filtering and the need for privacy.

Examples of environments where privacy is relevant and concrete examples where lack of privacy could have large costs

Methods for introducing security and privacy include differential privacy methods and encryption methods.

Differential privacy involves using statistical noise as security to make individual users' information cannot be deduced. Often requires a trusted aggregator, although secure aggregation methods exist. always requires noising result such that the outcome is not exact (a problem in localisation).

Encryption schemes involve formal indistinguishability proofs typically over bits or integers. They rely on computationally hard problems involving security parameters of a sufficiently large size; therefore the additional computational requirements of using encryption schemes should be pointed out and what this means in a real-time distributed sensor system. Continuing, explain public-key cryptography applicability to distributed systems; difference to symmetric schemes. Homomorphic encryption power and use case. Why FHE isn't used often, why additive partially homomorphic encryption is.

Advancements in function providing encryption schemes such as homomorphic encryption have also led to several other types of schemes which have found uses in signal processing. Private aggregation schemes allow the secure computation of the sum of encrypted values originating from different parties, leaking only the final result. When considering such multi-party encryption protocols, formal security definitions must now also incorporate the added dangers of colluding malicious parties, and lead to new notions of security. For example Aggregator Obliviousness (AO) is typically proven for private aggregation schemes, while alternatives such as Private Weighted Secure Aggregator Obliviousness (pWSAO) exist for other specific use-cases.

Another example of function providing encryption, and a generalisation of private aggregation, is called functional encryption (FE) and its distributed extension, multi-client functional encryption (MCFE), which allow the unencrypted result of an arbitrary function to be computed from encrypted inputs. General FE and MCFE are known to be quite computationally expensive (from meeting with ITI and student Johannes - need ref.) but alternatives providing only a subset of possibly computable function exist; for example, inner product encryption.

Several of the aforementioned encryption schemes have found uses in secure localisation, estimation, and control.

% 
% ##       #### ######## 
% ##        ##     ##    
% ##        ##     ##    
% ##        ##     ##    
% ##        ##     ##    
% ##        ##     ##    
% ######## ####    ##    
%

\subsection{Relevant Literature on Encrypted Localisation and Estimation}

Model-free localisation using homomorphic encryption examples include polygon thing, WSN examples which protect against adversaries but in the case of the WSN paper. don't preserve anchor privacy. Importantly, model-based filtering and localisation provide more accurate estimates and these are not applicable there.

Model-based estimation examples include Aristov paper (which requires a linear model, and a hierarchy of sensors), Farokhi paper (which requires the controller compute entirely in encrypted space and send input back to actuator - supporting only the cloud-as-a-service type architectures) and Alexandru paper (which implements a distributed control environment but requires a constant gain matrix K)

pWSAO achieved in Alexandru weighted aggregation, but requires redistributing keys at every timestep resulting in a costly operation, and a complicated communication protocol.

In addition to applying suitable encryption schemes to signal processing tasks, care must be taken when converting sensor output into an encryptable homomorphic format. As is the case with our proposed localisation method, real number sensor output does not trivially encode to integers such that the homomorphic properties provided by an additive encryption scheme over integers keep the underlying real numbers consistent. Methods for handling the encoding of real numbers such that they can be used in homomorphic encryption exist. Google bignum adds power but risks overflow and leaks exponents, Farokhi leaks no information but allows only a single multiplication (extendable to more but each further multiplication limits the real number size and increases the risk of overflow).

Briefly describe navigator scenario and our contributions

--

%Section Summary
% \ref{sec:problem_statement}
% \ref{sec:crypto_prelim}
% \ref{sec:lcao_definition}
% \ref{sec:our_scheme}
% \ref{sec:loc_prelim}
% \ref{sec:private_localisation}
% \ref{sec:results}
% \ref{sec:conclusion}
In Section \ref{sec:problem_statement} we introduce the localisation problem and the restrictions on involved parties. Section \ref{sec:crypto_prelim} summarises cryptographic preliminaries and Sections \ref{sec:lcao_definition} and \ref{sec:our_scheme} introduce a novel security notion of linear-combination aggregation obliviousness and an encryption scheme which satisfies it. Section \ref{sec:loc_prelim} summarises localisation preliminaries, before we introduce our private localisation method in Section \ref{sec:private_localisation}, using the introduced encryption scheme. In Section \ref{sec:results} we give and analyse simulation results, before concluding our work in Section \ref{sec:conclusion}.

% 
% ##    ##  #######  ########    ###    ######## ####  #######  ##    ## 
% ###   ## ##     ##    ##      ## ##      ##     ##  ##     ## ###   ## 
% ####  ## ##     ##    ##     ##   ##     ##     ##  ##     ## ####  ## 
% ## ## ## ##     ##    ##    ##     ##    ##     ##  ##     ## ## ## ## 
% ##  #### ##     ##    ##    #########    ##     ##  ##     ## ##  #### 
% ##   ### ##     ##    ##    ##     ##    ##     ##  ##     ## ##   ### 
% ##    ##  #######     ##    ##     ##    ##    ####  #######  ##    ## 
%

\subsection{Notation}
Notation. Reminder to include symbols for: rounding, concatenation, ``divides'', vector norm, expected value and variance, ordered list

% 
% 8888888b.                  888      
% 888   Y88b                 888      
% 888    888                 888      
% 888   d88P 888d888 .d88b.  88888b.  
% 8888888P"  888P"  d88""88b 888 "88b 
% 888        888    888  888 888  888 
% 888        888    Y88..88P 888 d88P 
% 888        888     "Y88P"  88888P"  
%                                     
%                                     
%                                     
%

\section{Problem Statement} \label{sec:problem_statement}
% Restate the scenario but more formally. Give a concrete example - plane and signal towers.

% Exact security guarantees we aim for, as well as the definitions for these guarantees (pWSAO and indistinguishability but in context of localisation as well). Note that learning only the sum in aggregation (as is normal in AO) would, in this case, tell the navigator the average location and measurements of all sensors, which is fine as it does not disclose any exact sensor.

% Passive attacks only from sensors to learn navigator position (Otherwise one could do some kind of attack that would send a fake measurement and note the change in its own measurements - possible this would give away the average of other sensors' measurements but unclear). Any largely incorrect inputs from sensors may also be detectable by comparison to alternative navigator onboard sensors (GPS etc.). Justify by saying sensors need to behave for localisation to work in the first place.

% Active attacks from navigator to find sensor location allowed, but assume that weights sent to all sensors are the same. In a wireless setting, all sensors would receive all broadcast weights anyway. While special hardware, which may support directional broadcasting or receiving, could be used to locate sensors individually this is beyond the scope of what is considered in our problem.

% Point out that learning the aggregation of sensor outputs, which contains measurement and location information also means that the average location and measurement of the sensors may be leaked, and is accepted as a part of the leakage as it is inferrable from the aggregation scheme and any functioning model-based localisation where measurements are not known

% Rough computational capabilities expected by parties

% Fixed sensor subsets of which only whole subsets can be used at once. Maybe a picture of what this might look like in a high level distributed localisation diagram. Should consider that this sub-grouping would also mean the leakage of the average sensor/measurement of each subset not all sensors at once. This should be considered when choosing sensor subsets and locations.

The localisation scenario we consider in this work is that of model-based self-navigation using range-only sensors. We consider localisation in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $w_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$, while the measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
is defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}_k) &= \left\lVert
        \begin{bmatrix}
            \rv{x}_k & \rv{y}_k
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}_k-s_{x,i})^2 + (\rv{y}_k-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn the estimated state $\rvec{x}_k$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$ or their measurements $\rv{z}_k$ at any time $k$.

We motivate these goals with an example. When considering aircraft navigation in the presence of privately-owned range-measuring towers, it is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown tower-owning parties. Similarly, tower locations may wish to be kept private from unidentified navigating aircraft. However, the additional safety to passengers, provided by accurate aircraft localisation, may be a goal of all those involved.

% 
%  ######     ###    ########     ###    ########  #### ##       #### ######## #### ########  ######  
% ##    ##   ## ##   ##     ##   ## ##   ##     ##  ##  ##        ##     ##     ##  ##       ##    ## 
% ##        ##   ##  ##     ##  ##   ##  ##     ##  ##  ##        ##     ##     ##  ##       ##       
% ##       ##     ## ########  ##     ## ########   ##  ##        ##     ##     ##  ######    ######  
% ##       ######### ##        ######### ##     ##  ##  ##        ##     ##     ##  ##             ## 
% ##    ## ##     ## ##        ##     ## ##     ##  ##  ##        ##     ##     ##  ##       ##    ## 
%  ######  ##     ## ##        ##     ## ########  #### ######## ####    ##    #### ########  ######  
%

\subsection{Participant Capabilities} \label{subsec:capabilities}
Before we can define concrete security requirements and a filtering protocol, we require some assumptions on the capabilities of the navigator and sensors in our problem scenario.
\begin{description}
    \item[Global navigator broadcast] Due to sensor location privacy and model non-linearity, bi-directional communication between navigator and sensors is required. We make the assumption that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within the range of a navigator broadcast. We loosely propose that this scenario can be handled by pre-defined sensor subsets during an initial step by a trusted party. This has been captured in Figure \ref{fig:sensor_subsets}.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is received equal by all sensors. This means the navigator may not send different information to individual sensors during a single timestep.

    In the context of wireless communication, the wide-spread use of cheap non-directional antennas make this assumption reasonable.
    \item[Honest-but-curious sensors] We adopt the honest-but-curious attacker model for all involved sensors, meaning they are assumed to follow the localisation procedure correctly but may store or use any gained sensitive information.

    Misbehaving sensors are a known problem in estimation theory, often requiring inconsistency or outlier detection mechanisms. As privacy-preserving implementations of such mechanisms require additional complications, we do not consider the scenario in this work.
    \item[Computational capabilities] The computational requirements for computing encryptions and homomorphic operations are typically more costly than their plaintext equivalents. We make the implicit assumption that all involved parties are computationally capable of computing the required encryption and filter operations.
\end{description}
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize,scale=0.9375]
    % Sizes
    \tikzmath{\srange = 3.54; \sradius = 0.04;}
    % Sensor locations (sensor 3 is shared)
    \coordinate (s1) at (1.4086,-4.5869);
    \coordinate (s2) at (4.0633,-5.9948);
    \coordinate (s3) at (3.1235,-3.2054);
    \coordinate (s4) at (4.5563,-1.6861);
    \coordinate (s5) at (5.9909,-3.5635);
    % Navigator location
    \coordinate (nav_loc) at (4,-4);
    % Bounding rectangle
    \coordinate (brect_start) at (0,0);
    \coordinate (brect_end) at (8.0, -8.0);

    % Gray range circles
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \draw[gray!50] (s1) circle (\srange);
        \draw[gray!50] (s2) circle (\srange);
        \draw[gray!50] (s3) circle (\srange);
        \draw[gray!50] (s4) circle (\srange);
        \draw[gray!50] (s5) circle (\srange);
    \end{scope}
    % Red subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[red!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Blue subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \fill[blue!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Magenta double overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[magenta!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Sensors
    \fill[black] (s1) circle (\sradius);
    \fill[black] (s2) circle (\sradius);
    \fill[black] (s3) circle (\sradius);
    \fill[black] (s4) circle (\sradius);
    \fill[black] (s5) circle (\sradius);
    % Bounding box
    \draw[gray] (brect_start) rectangle (brect_end);
    % Navigator
    \clip (brect_start) rectangle (brect_end);
    \pic[rotate=-40,xscale=0.22,yscale=0.3] at (nav_loc) {plane};
    \begin{scope}[shift={(nav_loc)},rotate=-40,xscale=0.22,yscale=0.3]
        % Trail
        \fill[bottom color=white, top color=gray] (-0.05, -1.5) rectangle (0.05, -50);
    \end{scope}
    % Sensor labels
    \node[label=1] at (s1) {};
    \node[label=2] at (s2) {};
    \node[label=3] at (s3) {};
    \node[label=4] at (s4) {};
    \node[label=5] at (s5) {};
\end{tikzpicture}
\caption{An example of sensor subsets; $\{1,2,3\}$ and $\{3,4,5\}$, and the areas where the navigator is in range of all sensors within a subset.}
\label{fig:sensor_subsets}
\end{figure}

% 
% ##       ########    ###    ##    ##    ###     ######   ######## 
% ##       ##         ## ##   ##   ##    ## ##   ##    ##  ##       
% ##       ##        ##   ##  ##  ##    ##   ##  ##        ##       
% ##       ######   ##     ## #####    ##     ## ##   #### ######   
% ##       ##       ######### ##  ##   ######### ##    ##  ##       
% ##       ##       ##     ## ##   ##  ##     ## ##    ##  ##       
% ######## ######## ##     ## ##    ## ##     ##  ######   ######## 
%

\subsection{Inherent Leakage}
Leakage of information will be discussed in more detail in Section \ref{subsec:leakage}, however, we stress that in any estimation problem where process and measurement models are known, knowledge of sequential state estimates naturally leads to the leakage of some measurement information. For this reason, we accept that average information about sensors may be leaked at each time step and instead show that individual sensor information remains private.

% 
%  .d8888b.                            888                 8888888b.                  888 d8b               
% d88P  Y88b                           888                 888   Y88b                 888 Y8P               
% 888    888                           888                 888    888                 888                   
% 888        888d888 888  888 88888b.  888888 .d88b.       888   d88P 888d888 .d88b.  888 888 88888b.d88b.  
% 888        888P"   888  888 888 "88b 888   d88""88b      8888888P"  888P"  d8P  Y8b 888 888 888 "888 "88b 
% 888    888 888     888  888 888  888 888   888  888      888        888    88888888 888 888 888  888  888 
% Y88b  d88P 888     Y88b 888 888 d88P Y88b. Y88..88P      888        888    Y8b.     888 888 888  888  888 
%  "Y8888P"  888      "Y88888 88888P"   "Y888 "Y88P"       888        888     "Y8888  888 888 888  888  888 
%                         888 888                                                                           
%                    Y8b d88P 888                                                                           
%                     "Y88P"  888                                                                           
%

\section{Cryptography Preliminaries} \label{sec:crypto_prelim}
When defining our system security requirements and encryption scheme, we will reference some existing cryptographic security notions, the additively homomorphic Paillier encryption scheme, and the Joye-Libert private aggregation scheme.

% 
%  ######  ########  ######  ##     ## ########  #### ######## ##    ## 
% ##    ## ##       ##    ## ##     ## ##     ##  ##     ##     ##  ##  
% ##       ##       ##       ##     ## ##     ##  ##     ##      ####   
%  ######  ######   ##       ##     ## ########   ##     ##       ##    
%       ## ##       ##       ##     ## ##   ##    ##     ##       ##    
% ##    ## ##       ##    ## ##     ## ##    ##   ##     ##       ##    
%  ######  ########  ######   #######  ##     ## ####    ##       ##    
%

\subsection{Security Notions}
The security of a cryptographic scheme is typically defined by a security \textit{game}, which captures both the desired privacy guarantees, as well as the capabilities of attackers \cite{katzIntroductionModernCryptography2008}. The typical security notion for a homomorphic encryption scheme is Indistinguishability under Chosen Plaintext Attack (IND-CPA) \cite{chaseSecurityHomomorphicEncryption2017}. 
\begin{defn}
An encryption scheme meets IND-CPA security if an attacker who can choose plaintext messages to be encrypted at will, gains no additional information about an unknown plaintext message when they learn only its encryption.

The formal security game for IND-CPA has been given in Appendix \ref{app:ind_cpa}.
\end{defn}
Private aggregation schemes aim for the security notion of Aggregator Obliviousness (AO) \cite{shiPrivacyPreservingAggregationTimeSeries2011}. 
\begin{defn}
An encryption scheme meets AO security if no colluding subset of participants \textit{excluding} the aggregator gains additional information about the remaining aggregation values given only their encryptions, while any colluding subset \textit{including} the aggregator learns only their sum.

The formal security game for AO has been given in Appendix \ref{app:ao}.
\end{defn}

% 
% ########     ###    #### ##       ##       #### ######## ########  
% ##     ##   ## ##    ##  ##       ##        ##  ##       ##     ## 
% ##     ##  ##   ##   ##  ##       ##        ##  ##       ##     ## 
% ########  ##     ##  ##  ##       ##        ##  ######   ########  
% ##        #########  ##  ##       ##        ##  ##       ##   ##   
% ##        ##     ##  ##  ##       ##        ##  ##       ##    ##  
% ##        ##     ## #### ######## ######## #### ######## ##     ## 
%

\subsection{Paillier Encryption Scheme} \label{subsec:paillier_scheme}
The Paillier encryption scheme \cite{paillierPublicKeyCryptosystemsBased1999} is an additively homomorphic encryption scheme which bases its security on the decisional composite residuosity assumption (DCRA) and meets the security notion of IND-CPA. Key generation of the Paillier scheme is performed by choosing two sufficiently large primes $p$ and $q$, and computing $N=pq$. A generator $g$ is also required for encryption, which is often set to $g=N+1$ when $p$ and $q$ are of equal bit length \cite{katzIntroductionModernCryptography2008}. The public key is defined by $(N, g)$ and secret key by $(p, q)$.

Encryption of a plaintext message $m \in \mathbb{Z}_N$, producing ciphertext $c \in \mathbb{Z}^{*}_{N^2}$, is computed by
\begin{equation}
    c = g^m r^N \pmod{N^2}
\end{equation}
for a randomly chosen $r \in \mathbb{Z}_{N}$. $r^N$ can be considered the noise term which hides the value $g^m \pmod{N^2}$, which due to the scheme construction, is an easily computable discrete logarithm. The decryption of a ciphertext is computed by
\begin{equation}
    m = \frac{L(c^\lambda\pmod{N^2})}{L(g^\lambda\pmod{N^2})} \pmod{N}
\end{equation}
where $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$.

In addition to encryption and decryption, the following homomorphic functions are provided by the Paillier scheme. $\forall m_1,m_2 \in \mathbb{Z}_N$,
\begin{align}
    \mathcal{D}(\mathcal{E}(m_1)\mathcal{E}(m_2) \hspace{-7pt} \pmod{N^2}) &= m_1+m_2 \hspace{-7pt} \pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)g^{m_2} \hspace{-7pt} \pmod{N^2}) &= m_1+m_2\hspace{-7pt}\pmod{N} \\
    \mathcal{D}(\mathcal{E}(m_1)^{m_2} \hspace{-7pt} \pmod{N^2}) &= m_1m_2 \hspace{-7pt} \pmod{N}\,. \label{eqn:paillier_hom_mult}
\end{align}

% 
%       ##         ##    ## 
%       ##          ##  ##  
%       ##           ####   
%       ## #######    ##    
% ##    ##            ##    
% ##    ##            ##    
%  ######             ##    
%

\subsection{Joye-Libert Private Aggregation Scheme} \label{subsec:joye_libert_scheme}
The Joye-Libert private aggregation scheme \cite{joyeScalableSchemePrivacyPreserving2013} is a scheme defined on time-series data and meets the security notion of AO. Similarly to the Paillier scheme, it bases its security on the DCRA. A notable difference to a public-key encryption scheme is the need for a trusted party to perform an initial key generation and distribution step.

Key generation is computed by choosing two equal length and sufficiently large primes $p$ and $q$, and computing $N=pq$. Additionally, hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$ is defined, and the public key is set to $(N, H)$. $n$ private keys are generated by choosing $sk_i,\,1\leq i\leq n$ uniformly from $\mathbb{Z}_{N^2}$ and distributing them to all users, while the last key is set as
\begin{equation*}
    sk_0 = -\sum^{n}_{i=1}sk_i \pmod{N^2}\,,
\end{equation*}
and sent to the aggregator.

Encryption of plaintext $m^{(t)}_{i} \in \mathbb{Z}_N$ to ciphertext $c^{(t)}_{i} \in \mathbb{Z}_{N^2}$ at time $t$ is computed by user $i$ as
\begin{equation}
    c^{(t)}_{i} = (N+1)^{m^{(t)}_{i}} H(t)^{sk_i} \pmod{N^2}\,,
\end{equation}
where the $H(t)^{sk_i}$ can be considered the noise term which hides the again easily computable discrete logarithm $g^{m^{(t)}_{i}} \pmod{N^2}$, where $g=N+1$.

When all encryptions $c^{(t)}_{i},\,1\leq i \leq n$ are sent to the aggregator, private summation and decryption are computed by the functions
\begin{equation}
    c^{(t)} = H(t)^{sk_0}\prod^{n}_{i=1}c^{(t)}_{i} \pmod{N^2}
\end{equation}
and
\begin{equation}
    \sum^{n}_{i=1}m^{(t)}_{i} = \frac{c^{(t)}-1}{N}\,. \label{eqn:agg_decryption}
\end{equation}
Correctness follows from $\sum^{n}_{i=0}sk_i = 0$, and thus
\begin{equation*}
    \begin{split}
        &H(t)^{sk_0}\prod^{n}_{i=1}c_{i,t} \pmod{N^2} \\
        \equiv &H(t)^{sk_0}\prod^{n}_{i=1}(N+1)^{m_{i,t}} H(t)^{sk_i} \pmod{N^2} \\
        \equiv &H(t)^{\sum^n_{j=0}sk_j} \prod^{n}_{i=1}g^{m_{i,t}} \pmod{N^2} \\
        \equiv &(N+1)^{\sum^n_{i=1}m_{i,t}} \pmod{N^2}
    \end{split}
\end{equation*}
removing all noise terms.

% 
%  .d8888b.                            888                 8888888b.                  888      
% d88P  Y88b                           888                 888   Y88b                 888      
% 888    888                           888                 888    888                 888      
% 888        888d888 888  888 88888b.  888888 .d88b.       888   d88P 888d888 .d88b.  88888b.  
% 888        888P"   888  888 888 "88b 888   d88""88b      8888888P"  888P"  d88""88b 888 "88b 
% 888    888 888     888  888 888  888 888   888  888      888        888    888  888 888  888 
% Y88b  d88P 888     Y88b 888 888 d88P Y88b. Y88..88P      888        888    Y88..88P 888 d88P 
%  "Y8888P"  888      "Y88888 88888P"   "Y888 "Y88P"       888        888     "Y88P"  88888P"  
%                         888 888                                                              
%                    Y8b d88P 888                                                              
%                     "Y88P"  888                                                              
%

\section{Private Linear-Combination Aggregation} \label{sec:lcao_definition}
For achieving the goal of private localisation defined in Section \ref{sec:problem_statement} we will require a multi-party protocol for computing linear combinations of weights, and their subsequent aggregation. In the context of navigation, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are first broadcast by the navigator, linear combinations $y^{(t)}_i=\sum^m_{j=1}x_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator, at every time step $t$. This has been summarized in Figure \ref{fig:agg_steps}. In Section \ref{sec:private_localisation} we will show how this protocol can be sufficient to compute measurement covariances and measurement vectors, and update the Extended Information Filter.

\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize,scale=0.75]
    % Step 1
    \node at (5,22.5) {1. Broadcast and linear combinations step};
    % Navigator
    \node at (1.25,19.25) {Navigator};
    \fill (1.25,18.25) [blue!40] ellipse (0.5 and 0.5);
    % Sensors
    \node at (7.5,21.75) {Sensor $1$};
    \fill [red!40] (5.875,19.75) rectangle (9.5,21.25);
    \node at (7.75,20.5) {$\displaystyle y_1^{(t)} = \sum^m_{j=1}x_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (7.5,14.75) {Sensor $n$};
    \fill [red!40] (5.875,15.25) rectangle (9.5,16.75);
    \node at (7.75,16) {$\displaystyle y_n^{(t)} = \sum^m_{j=1}x_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (7.5,18.75) circle (0.05);
    \fill [black] (7.5,17.75) circle (0.05);
    \fill [black] (7.5,18.25) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(1.75,18.5) (5.75,20.25)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(1.75,18) (5.75,16.25)};
    \node at (3.5,20.25) {$(\omega_1^{(t)},\dots ,\omega_m^{(t)})$};
    \node at (3.5,16.25) {$(\omega_1^{(t)},\dots ,\omega_m^{(t)})$};

    % Step 2
    \node at (5,13.25) {2. Aggregation step};
    % Navigator
    \node at (3.25,10.75) {Navigator};
    \fill [blue!40] (0.5,8.75) rectangle (5.625,10.25);
    \node at (3,9.5) {$\displaystyle \sum^{n}_{i=1}\sum^{m}_{j=1} x_{i,j}^{(t)}\omega_j^{(t)} = \sum^n_{i=1}y^{(t)}_{i}$};
    % Sensors
    \node at (8.75,12.5) {Sensor $1$};
    \fill  (8.75,7.5) [red!40] ellipse (0.5 and 0.5);
    \node at (8.75,6.5) {Sensor $n$};
    \fill  (8.75,11.5) [red!40] ellipse (0.5 and 0.5);
    \fill [black] (8.75,10) circle (0.05);
    \fill [black] (8.75,9) circle (0.05);
    \fill [black] (8.75,9.5) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(8.25,11.25) (5.75,10)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(8.25,7.75) (5.75,9)};
    \node at (6.75,11.5) {$y_1^{(t)}$};
    \node at (6.75,7.5) {$y_n^{(t)}$};

    % Bounding rectangles
    \draw [gray] (0,23.25) rectangle (10,14.25);
    \draw [gray] (0,14) rectangle (10,5.5);
\end{tikzpicture}
\vspace{\baselineskip}
\caption{Required linear-combination aggregation steps at time $t$.}
\label{fig:agg_steps}
\end{figure}

>>>>

\begin{tikzpicture}[font=\footnotesize]
% Bounding box
\draw [gray] (0,-0.25) rectangle (8,7.5);
